---
title: "p8105_hw2_wy2369"
date: 10/3/2022
author: Wenjing Yang
output: github_document
---

# Problem 1

### Import and clean data

```{r load_libraries, message = FALSE}
library(tidyverse)
library(readxl)
```

The process begins with data import, updates variable names, and selects the columns that will be used in later parts fo this problem. Update `entry` from `yes` / `no` to a logical variable using `ifelse`.

As part of data import, we specify that `Route` columns 8-11 should be character for consistency with 1-7.

```{r}
trans_ent = 
  read_csv(
    "./hw2data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

### Describe the dataset in a short paragraph.

Read data from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`, and clean it using `janitor::clean_names()`.The dataset contains `line, station_name, station_latitude, station_longitude, route, entry, exit_only, vending, entrance_type, and ada` variables. The resulting dataset has **`r nrow(trans_ent)`** rows and  **`r ncol(trans_ent)`** columns. 


These data are not "tidy" because route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format. 

### Explain how many distinct stations are there.

Use `distinct()` to select station name and line, and then obtain all unique combinations. As a result, the number of rows in this dataset is the number of unique stations. There are **465** distinct stations. 

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```


### Explain how many stations are ADA compliant. 

Use `filter()` as an initial step to find how many stations are ADA cpmpliant. This produces a dataframe in which the number of rows is the number of ADA compliant stations. Hence, **84** stations are ADA compliant.

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

### Compute the proportion of station entrances / exits without vending allow entrance.

First exclude station entrances that do not allow vending. Then, use `entry` variable -- this logical. Taking the mean will produce the desired proportion, which is **0.3770492**.

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

### Compute the stations that serve the A train and how many A train stations are ADA compliant. 

Firstly, tidy the data as alluded to previously and convert `route` from wide to long format by using `pivot_longer()`. After this step, we can use tools from previous parts of the question. We get **60** stations that serve A train and **17** of these stations are ADA compliant.

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```



# Problem 2

### Import, clean, and organize **Mr. Trash Wheel** sheet.

Use code `read_excel` to read, and `janitor::clean_names()` to clean data from `Trash Wheel Collection Data.xlsx`. Omit non-data rows and columns, and then round the number of sports balls to the nearest integer. After that, convert the result to an integer variable by using `as.integer`.

```{r}
mr_trash_df = 
  read_excel(
    "./hw2data/Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N550") %>% 
  janitor::clean_names() %>%
  drop_na (c(dumpster)) %>%
  mutate(sports_balls = round(sports_balls,digits = 0) )%>%
  mutate(sports_balls = as.integer(sports_balls) ,data_source = "mr_trash_wheel",year = as.integer(year))

```

### Import, clean, and organize **Professor Trash Wheel** sheet.

Use a similar process to organize the data from Professor Trash Wheel sheet.

```{r}
prof_trash_df = 
  read_excel(
    "./hw2data/Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M97") %>% 
  janitor::clean_names() %>%
  drop_na (c(dumpster)) %>%
  mutate(data_source = "prof_trash_wheel")
```

  
### Combine the two datasets.

Use code `bind_rows` to combine `mr_trash_df` and `prof_trash_df` , and get `trash_tidy` dataset.
```{r}
trash_tidy = 
  bind_rows(mr_trash_df,prof_trash_df) %>%
  janitor::clean_names() 

trash_tidy
```


### Write a short description and answer questions.

In trash_tidy dataset, there are 15 key variables, for example month, date, weight_tons, and glass_bottles. The data frame has **`r nrow(trash_tidy)`** rows and **`r ncol(trash_tidy)`** columns. 

```{r include = TRUE}
weight_proTrash = 
  trash_tidy %>%
  filter(data_source == "prof_trash_wheel") 
  sum(weight_proTrash$weight_tons)
```
The total weight of trash collected by Professor Trash Wheel is **190.12**.


```{r}
year_2020 = 
  trash_tidy %>%
  filter(data_source == "mr_trash_wheel", year == 2020)
  sum(year_2020$sports_balls)
```
The total number of sports balls collected by Mr. Trash Wheel in 2020 is **856**.


# Problem 3

### Read and clean data from `fivethirtyeight` datasets.

Use this code chunk to read and organize `pols-month.csv` dataset. Use `separate()` to break up the variable **mon** into integer variables year, month, and day; Use `month.abb` to replace month number with month name. Create a president variable taking values "DEM" and "GOP", and remove prez_dem, prez_gop and the day variable in the dataset.

```{r message = FALSE}
pols_month =
  read_csv("./fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year","month","day") ) %>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.abb[month],
    president = ifelse( prez_dem == 1, "DEM","GOP" )) %>%
  select(-prez_dem,-prez_gop,-day)
  
pols_month
```

Use this code chunk to read and organize `snp.csv` dataset, and arrange according to year and month, and organize so that year and month are the leading columns.

```{r message = FALSE}
snp_df =
  read_csv("./fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  mutate(date = as.Date(date, "%m/%d/%y")) %>%
  separate(date, into = c("year","month","day") ) %>%
  mutate(
    year = as.integer(year),
    year = ifelse(year > 2049, year-100, year),
    month = as.integer(month),
    day = as.integer(day)) %>%
  arrange(year, month) %>%
  mutate(month = month.abb[month])  %>%
  select(-day)

snp_df
```


Use this code chunk to read and organize `unemployment.csv` dataset. This process will involve switching from “wide” to “long” format using `pivot_longer()`.

```{r message = FALSE}
unemploy_df =
  read_csv("./fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  mutate(
    year = as.integer(year)) %>%
  pivot_longer(
    jan : dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) %>%
  mutate(month = str_to_title(month))
  
unemploy_df
```



### Join the datasets. 

Merging `snp_df` into `pols_month`, and merging `unemploy_df` into the result using this code chunk. 

```{r}
snp_pols_tidy = 
  left_join(pols_month, snp_df, by = c("year","month"))

snp_pols_unemlop_tidy = 
  left_join(snp_pols_tidy, unemploy_df,by = c("year","month"))

snp_pols_unemlop_tidy
```

### Explain the resulting dataset. 

For `pols_month`, it has `r nrow(pols_month)` rows and `r ncol(pols_month)` columns. There are **9** distinct variables which include year, month, gov_gop, sen_gop, rep_gop, gov_dem, sen_demm, rep_dem, and president. The range of year in the dataset is **from 1947 to 2015**.

For `snp_df`, it has `r nrow(snp_df)` rows and `r ncol(snp_df)` columns. There are **3** distinct variables which include year, month, and close in the dataset. The range of year is **from 1950 to 2015**.

For `unemploy_df`, it has `r nrow(unemploy_df)` rows and `r ncol(unemploy_df)` columns. There are **3** distinct variables which include year, month, and unemployment_rate in the dataset. The range of year is **from 1948 to 2015**.


Use code `left_join()` to combine these datasets. In the resulting dataset `snp_pols_unemlop_tidy`, it has `r nrow(snp_pols_unemlop_tidy)` rows and `r ncol(snp_pols_unemlop_tidy)` columns, and **11** key variables which are year, month, gov_gop, sen_gop, rep_gop, gov_dem, sen_demm, rep_dem, president, close, and unemployment_rate. The range of year in the resulting dataset is **from 1947 to 2015**.
