---
title: "p8105_hw2_wy2369"
date: 10/3/2022
author: Wenjing Yang
output: github_document
---

# Problem 1

### Import and clean data

```{r load_libraries, message = FALSE}
library(tidyverse)
library(readxl)
```

The process begins with data import, updates variable names, and selects the columns that will be used in later parts fo this problem. Update `entry` from `yes` / `no` to a logical variable using `ifelse`.

As part of data import, we specify that `Route` columns 8-11 should be character for consistency with 1-7.

```{r}
trans_ent = 
  read_csv(
    "./hw2data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

### Describe the dataset in a short paragraph.

Read data from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`, and clean it using `janitor::clean_names()`.The dataset contains `line, station_name, station_latitude, station_longitude, route, entry, exit_only, vending, entrance_type, and ada` variables. The resulting dataset has **`r nrow(trans_ent)`** rows and  **`r ncol(trans_ent)`** columns. 


These data are not "tidy" because route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format. 

### Explain how many distinct stations are there.

Use `distinct()` to select station name and line, and then obtain all unique combinations. As a result, the number of rows in this dataset is the number of unique stations. There are **465** distinct stations. 

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```


### Explain how many stations are ADA compliant. 

Use `filter()` as an initial step to find how many stations are ADA cpmpliant. This produces a dataframe in which the number of rows is the number of ADA compliant stations. Hence, **84** stations are ADA compliant.

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

### Compute the proportion of station entrances / exits without vending allow entrance.

First exclude station entrances that do not allow vending. Then, use `entry` variable -- this logical. Taking the mean will produce the desired proportion, which is **0.3770492**.

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

### Compute the stations that serve the A train and how many A train stations are ADA compliant. 

Firstly, tidy the data as alluded to previously and convert `route` from wide to long format by using `pivot_longer()`. After this step, we can use tools from previous parts of the question. We get **60** stations that serve A train and **17** of these stations are ADA compliant.

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```



# Problem 2

### Import, clean, and organize **Mr. Trash Wheel** sheet.

Use code `read_excel` to read, and `janitor::clean_names()` to clean data from `Trash Wheel Collection Data.xlsx`. Omit non-data rows and columns, and then round the number of sports balls to the nearest integer. After that, convert the result to an integer variable by using `as.integer`.

```{r}
mrtrash_df = 
  read_excel(
    "./hw2data/Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N550") %>% 
  janitor::clean_names()


```

### Import, clean, and organize **Professor Trash Wheel** sheet.

Use 

  drop_na (c("Dumpster") )
  mutate(sports_balls = round(sports_balls,digits = 0) 






# Problem 3


